\clearpage{\pagestyle{empty}\cleardoublepage}
\begin{thebibliography}{3}
\addcontentsline{toc}{chapter}{Bibliografia}

\bibitem{specsSOAP} N.Mitra et al. (2003), \emph{"SOAP Version 1.2"} recommendation del W3C - \newline http://www.w3.org/TR/soap12-part0/

\bibitem{specsWSIBP} K.Ballinger, D.Ehnebuske et al. (2004), \emph{"Basic Profile Version 1.1"} specifica del gruppo WS-I - \newline http://www.ws-i.org/Profiles/BasicProfile-1.1.html

\bibitem{specsWSDL} D.Booth, C.K.Liu et al. (2004), \emph{"Web Services Description Language (WSDL) Version 2.0"} working draft del W3C - \newline http://www.w3.org/TR/wsdl20-primer/

\bibitem{specsWSS1} A.Nadalin, C.Kaler et al. (2004), \emph{"Web Services Security: SOAP Message Security"} standard OASIS - \newline http://docs.oasis-open.org/wss/2004/01/oasis-200401-wss-soap-message-security-1.0.pdf

\bibitem{specsWSS2} A.Nadalin, P.Griffin et al. (2004), \emph{"Web Services Security: UsernameToken Profile"} standard OASIS - \newline http://docs.oasis-open.org/wss/2004/01/oasis-200401-wss-username-token-profile-1.0.pdf

\bibitem{specsWSS3} P.Hallam-Baker, C.Kaler et al. (2004), \emph{"Web Services Security: X.509 Certificate Token Profile"} standard OASIS - \newline http://docs.oasis-open.org/wss/2004/01/oasis-200401-wss-x509-token-profile-1.0.pdf

\bibitem{specsWCAG} W. Chisholm, G.Vanderheiden et al. (1999), \emph{"Web Content Accessibility Guidelines 1.0"} recommendation del W3C - \newline http://www.w3.org/TR/WAI-WEBCONTENT/


\bibitem{bookPWF} J.Zeldman (2003), \emph{"Progettare il web del futuro"} ed. Pearson Education Italia

\bibitem{bookXPEplained} K.Beck (2000), \emph{"eXtreme Programming explained"} ed. Addison-Wesley

\bibitem{bookRefactoring} M.Fowler (1999), \emph{"Refactoring - Improving the design of existing code"} ed. Addison-Wesley

\bibitem{bookTDD} K.Beck (2002), \emph{"Test Driven Development"} ed.Addison-Wesley

\bibitem{bookUserStories} M.Cohn (2004) "User Stories Applied: For Agile Software Development" ed. Addison-Wesley

\bibitem{bookAgileSD} A.Cockburn (2002), \emph{"Agile Software Development"} ed. Pearson Education

\bibitem{bookJ2EESec} P.Kumar (2004), \emph{"J2ee Security"} ed. Prentice Hall

\bibitem{bookCVS} K.Fogel, M.Bar (2003) \emph{"Open Source Development with CVS, 3rd Edition"} ed. O'Reilly, distribuito con licenza FDL e disponibile liberamente da \newline http://cvsbook.red-bean.com/

\bibitem{artSoc} M.P.Papazoglou, D. Georgakopoulus (2003), \emph{"Service-Oriented Computing"} introduzione a \emph{Communications of the ACM} vol.46 n.10-Ottobre

\bibitem{tesiGerminale} M.Germinale (2003), \emph{"Progetto di applicazioni basate su Web Services"} tesi di laurea in Ingegneria Informatica presso l'Università di Bologna

\bibitem{tesiVigno} M.Vignoli (2004), \emph{"Analisi e progettazione di un servizio di tipo ICT per la didattica"} - tesi di laurea in Ingegneria Gestionale presso l'Università di Bologna

\bibitem{CatEBaz}  E.S.Raymond (1997), \emph{"The Cathedral and the Bazaar"}  \newline http://www.catb.org/~esr/writings/cathedral-bazaar/

\bibitem{HackAndRef}  E.S.Raymond (2003), \emph{"Hacking and refactoring: discovering the obvious"} \newline http://www.artima.com/weblogs/viewpost.jsp?thread=5342

\bibitem{linkWSresearch} J.Hurwitz (2004) \emph{Web Services Primary Research Opportunity Study} \newline http://www.hurwitz.com/prostudies/webservices.htm

\bibitem{linkWSvsDO} D.Orchard (2004), \emph{"Web services == or != distributed objects?"} \newline http://www.pacificspirit.com/blog/2004/01/12/web\_services\_or\_distributed\_objects

\bibitem{linkWSinsec} B.Schneier (2001), \emph{"SOAP insecurity"} \newline http://www.counterpane.com/crypto-gram-0006.html

\bibitem{linkXsdExt} D.Orchard (2003). \emph{"Examining wildcards for versioning"}  \newline http://www.pacificspirit.com/Authoring/Compatibility/Examining-\newline ElementWildcardSiblings.html

\bibitem{linkAgileManifesto} K.Beck,A.Cockburn,M.Fowler et. al. (2001) \emph{"Manifesto for Agile Software Development"} \newline http://agilemanifesto.org/

\bibitem{linkRolesFowler} M.Fowler (1997) \emph{"Dealing with roles"} \newline http://www.martinfowler.com/apsupp/roles.pdf

\bibitem{linkAppSecurity} J.Yoder e J.Barcalow (1998) \emph{Architectural Patterns for Enabling Application Security} \newline http://citeseer.nj.nec.com/yoder98architectural.html

\bibitem{linkJAXRPC} Java Community Process (2003) \emph{JSR 101: Java APIs for XML based RPC} \newline http://www.jcp.org/en/jsr/detail?id=101

\bibitem{linkJAXRPCSun} Sun (2003) \emph{Java(TM) API for XML-Based RPC Documentation} \newline http://developers.sun.com

\bibitem{linkPHP} Apache Foundation \emph{"Php: Hypertext Preprocessor"} \newline http://www.php.net

\bibitem{linkPHPeSSQLInjection} Tomas Oertli (2000) \emph{"Secure PHP Programming"} \newline www.zend.com/zend/art/art-oertli.php

\bibitem{linkFCPattern} \emph{"Wiki Web Application Component Toolkit - Front Controller"} \newline http://wact.sourceforge.net/index.php/FrontController

\bibitem{linkToolbox} J.B. Rainsberger (2001), \emph{"Toolbox: Use your singletons wisely"} \newline http://www-106.ibm.com/developerworks/webservices/library/co-single.html

\bibitem{PHPeWS} I. Bartolini (2003), \emph{"La sicurezza nelle applicazioni Web con PHP"} \newline http://www.phpday.it/download/2003-10-24/2\_PHP\_Web\_Security.pdf oppure http://www.brainetwork.net/homepage/txs/

\bibitem{PHPeDBAL} I. Bartolini (2003), \emph{"DataBase Abstraction Layers
in PHP"} \newline http://www.brainetwork.net/homepage/si2/

\bibitem{linkPearDbDoc} S.Bakken, T.V.V.Cox, D.Convissor (2004), \emph{"PEAR::DB Documentation"} \newline http://pear.php.net/package/DB

\bibitem{bookStat} A.Forcina, E.Stanghellini (2002) \emph{"Elementi di Statistica"} ed. Morlacchi Perugia

\end{thebibliography} 
